---
title: "Snow and Irradiance Effects on PV Performance (Beta Regression)"
author: Lucian Scher
date: "12/4/25"
format:
  html:
    toc: true
    toc-title: "Contents"
execute:
  echo: true
  warning: false
  message: false
---

## Purpose 

This blog post builds on Jackson & Gunda's (2021) machine learning analysis of United States solar systems, who showed how weather events, especially snow, can reduce photovoltaic (PV) performance measured as PR (performance ratio). This analysis uses a simple beta regression to link daily performance ratio (PR) to cumulative snow depth, low-irradiance conditions, and basic solar plant age/region effects.

### Reference:
Jackson, N. D., & Gunda, T. (2021). Evaluation of extreme weather impacts on utility-scale photovoltaic plant performance in the United States. Applied Energy, 302, 117508. https://doi.org/10.1016/j.apenergy.2021.117508

## Hypothesis

After accounting for plant age and regional differences, cumulative snow depth
and low-irradiance will reduce the performance ratio (PR) of photovoltaics.

## Setup

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(betareg)
library(broom)
library(kableExtra)
library(dagitty)

# Load data
oedi_data <- read_excel("data/oedi_data.xlsx", sheet = "data") |>
  clean_names()

```

## Data wrangling 

PR must be strictly between 0 and 1 for beta regression. We drop rows with missing
inputs and slightly reduce 1s.

```{r}
notzero <- 1e-4 # Basically 0

beta_df <- oedi_data |>
  mutate(
    region = factor(noaa_clim_region),
    date = as.Date(date),
    pr_beta = pmin(pmax(pr, notzero), 1 - notzero)
  ) |>
  drop_na(region, date, pr_beta, cumulative_snow_mm, low_irradiation, plant_age_months) |>
  select(region, date, pr_beta, pr, cumulative_snow_mm, low_irradiation, plant_age_months)

glimpse(beta_df)
```

## DAG (drivers we adjust for)

```{r}
pr_dag <- dagitty("
dag {
  Region -> Snow
  Region -> LowIrr
  Region -> PR
  PlantAge -> PR
  Snow -> PR
  LowIrr -> PR
}
")

plot(pr_dag)
```

Region can affect weather and baseline PR, snow and low irradiance directly affect PR,
plant age captures aging machinery effects.

## Model concept

Beta regression with logit link:

logit(μ) = β0 + β1·Snow + β2·LowIrr + β3·Age + Region effects

PR ~ Beta(μ, φ)

## Fit beta regression

```{r}
beta_fit <- betareg(
  pr_beta ~ cumulative_snow_mm + low_irradiation + plant_age_months + region,
  data = beta_df,
  link = "logit"
)

beta_fit
```

## Coefficients (tidy)

```{r}
coef_tbl <- tidy(beta_fit) |>
  select(term, estimate, std.error, statistic, p.value)

coef_tbl |>
  kable(digits = 3, caption = "Beta regression on PR") |>
  kable_styling(full_width = FALSE)
```

The most impactful predictor is low irradiation, which is associated with a large, highly significant decrease (a logit shift of -1.084) in the logit of PR. Plant age also significantly drives PR lower as it increases ($\text{logit}(\text{PR})$ decreases by $0.004$ per month). For comparison, the amount of cumulative snow is not a statistically significant predictor of PR in this model ($p=0.309$).

## Effect of snow depth (holding other factors)

```{r}
region_ref <- beta_df |>
  count(region, sort = TRUE) |>
  slice(1) |>
  pull(region)

snow_grid <- tibble(
  cumulative_snow_mm = seq(0, quantile(beta_df$cumulative_snow_mm, 0.95, na.rm = TRUE), length.out = 100),
  low_irradiation = 0,
  plant_age_months = median(beta_df$plant_age_months, na.rm = TRUE),
  region = factor(region_ref, levels = levels(beta_df$region))
)

snow_grid <- snow_grid |>
  mutate(pred_pr = predict(beta_fit, newdata = snow_grid, type = "response"))

snow_grid |>
  ggplot(aes(cumulative_snow_mm, pred_pr)) +
  geom_line(color = "purple", linewidth = 1) +
  labs(
    title = "Predicted PR vs. snow depth (typical irradiance)",
    x = "Cumulative snow depth (mm)",
    y = "Predicted performance ratio"
  ) +
  theme_minimal()
```

## Results

Our Beta regression model partially supported the hypothesis that cumulative snow depth and low-irradiance would reduce photovoltaic performance. Low-irradiance days were significantly associated with a reduction in the log-odds of the Performance Ratio (PR). While the coefficients for cumulative snow depth also suggested a negative impact on PR after accounting for plant age and regional baselines, this effect was not statistically significant. Therefore, the primary driver supporting our hypothesis was the significant negative impact of low-irradiance conditions on photovoltaic performance.

